The Role of Context in Spatial Region Identification

### 2013-11-15 scenarios/cdsr/trunk/papers/ACS2013/main.tex
jk: fixed some typos

\begin{abstract}
Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.  Consider an agent asked to remain ``in the safe zone'' during a military engagement.  The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.  We call these types of regions \textit{context-dependent spatial regions} (CDSRs) due to the context-dependent nature of their extent.  We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.  We report the results of a human study conducted to explore how people interpret three types of CDSRs and discuss the implications for developing cognitive systems with this ability.
\end{abstract}

@@

Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tieds to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.  We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.  To explore these claims, we present the results of a human study in which people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider the applicability of approaches from cognitive systems and robotics.  We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.  Current approaches have a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

\subsection{Task 2 - Rating Points} 
In task two, the subject was asked to rank three points on a five point \textit{Likert scale} how well the location matched the given CDSR, or select not applicable.  The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textit{sweetspot}, was located near the expected prototypical center of the CDSR. In the classroom domain, the three points were located along the horizontal line that bisected the stimulus with P1 positioned on the left of the stimulus, P2 in the center and P3 on the right of the stimulus.  This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.  In the geography situations, P1 was positioned in the top left corner of the stimulus, P2 was the expected \textit{sweetspot} and was located within the expected bottleneck region for both stimulus and P3 was located in the bottom right corner of the stimulus.  In the military engagement domain the three test points were located on the vertical line that bisected the stimulus: P1 was located near the enemy tanks, P3 was located with all friendly units between it and the enemy tanks and P2 was located near the center of the image.  The locations are shown in Figure \ref{fig:locations}.
%Kate: to save some space, maybe we could talk generally about location placement and only give specifics for one set of stimuli?  Just a though

@@

Whilst spatial reasoning has been an active topic in AI and robotics for decades, the properties we have demonstrated for CDSRs (i.e. that they vary systematically with the function and configuration of objects, plus the agents interacting with them) are not addressed by previous work, and thus it is not possible to take existing work and build a cognitive system capable of reasoning about, or communicating with, CDSRs. The remainder of this section reviews related previous work with reference to the requirements of modelling CDSRs.

Regions are central to qualiative spatial reasoning (QSR)~\cite{Cohn:2001}. Regions can either have crisp boundaries, usually defined by geometric shapes (e.g. the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g. the `egg-yolk model'~\cite{Cohn96b}) or by thresholding an underlying spatially-mapped function (e.g. those created by potential fields~\cite{brenneretal07ijcai} or by Gaussian kernels~\cite{burbridge-dearden12}). The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points seem unlikely to model the regions with sufficient flexibility.

>>>>

\begin{abstract}
Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.  Consider an agent asked to remain ``in the safe zone'' during a military engagement.  The size, shape, and location of this area depends on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.  We call these types of regions \textit{context-dependent spatial regions} (CDSRs) due to the context-dependent nature of their extent.  We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.  We report the results of a human study conducted to explore how people interpret three types of CDSRs and discuss the implications for developing cognitive systems with this ability.
\end{abstract}

@@

Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tied to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''.  We claim that (1) people can effectively communicate using these regions and (2) the size, location and shape of these regions change with the context.  To explore these claims, we present the results of a human study in which people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider the applicability of approaches from cognitive systems and robotics.  We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.  Current approaches have a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

\subsection{Task 2 - Rating Points} 
In task two, the subject was asked to rank three points on a five point \textit{Likert scale} with respect to how well the location of the point matched the given CDSR, or select not applicable.  The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textit{sweetspot}, was located near the expected prototypical center of the CDSR. In the classroom domain, the three points were located along the horizontal line that bisected the stimulus with P1 positioned on the left of the stimulus, P2 in the center and P3 on the right of the stimulus.  This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.  In the geography situations, P1 was positioned in the top left corner of the stimulus, P2 was the expected \textit{sweetspot} and was located within the expected bottleneck region for both stimulus and P3 was located in the bottom right corner of the stimulus.  In the military engagement domain the three test points were located on the vertical line that bisected the stimulus: P1 was located near the enemy tanks, P3 was located with all friendly units between it and the enemy tanks and P2 was located near the center of the image.  The locations are shown in Figure \ref{fig:locations}.
%Kate: to save some space, maybe we could talk generally about location placement and only give specifics for one set of stimuli?  Just a though

@@

Whilst spatial reasoning has been an active topic in AI and robotics for decades, the properties we have demonstrated for CDSRs (i.e. that they vary systematically with the function and configuration of objects, plus the agents interacting with them) are not addressed by previous work.  Thus it is not possible to take existing work and build a cognitive system capable of reasoning about, or communicating with, CDSRs.  The remainder of this section reviews related previous work with reference to the requirements of modelling CDSRs.

Regions are central to qualiative spatial reasoning (QSR)~\cite{Cohn:2001}. Regions can either have crisp boundaries, usually defined by geometric shapes (e.g. the bounding boxes of visually tracked objects~\cite{SridharCohn:10}), or vague boundaries created qualitatively by multiple crisp regions (e.g. the `egg-yolk model'~\cite{Cohn96b}) or by thresholding an underlying spatially-mapped function (e.g. those created by potential fields~\cite{brenneretal07ijcai,kelleher/costello:09} or by Gaussian kernels~\cite{burbridge-dearden12}). The fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points seem unlikely to model the regions with sufficient flexibility.
